---
title: "Human Activity Recognition - practical machine learning"
author: "Junho Song"
date: "26 Oct, 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document: default
---
<br><br><br><br><br>


# Background  
In this project, as prctical machin learning course of cousera we are going to anlyize HAR(human activity recognition) data. The folowing is overview of data sets.  

=====
 In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).
 
Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions:
1)exactly according to the specification (Class A),
2)throwing the elbows to the front (Class B),
3)lifting the dumbbell only halfway (Class C),
4)lowering the dumbbell only halfway (Class D) and
5)throwing the hips to the front (Class E).
=====

# Data PreProcessing 

Set data folder
```{r}
filePath <- "/Users/uxLab/Box Sync/PROJECTS/MOOC/Johns Hopkins - Data Science/8 Practical Machine Learning/project/data/"
```

Check the original data set
```{r}
fileNameTraining <- "pml-training.csv"
fileNameTesting <- "pml-testing.csv"

fileUrlTraining <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileUrlTesting <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

Load data sets
```{r}
fileTraining <- paste0(filePath, fileNameTraining)
fileTesting <- paste0(filePath, fileNameTesting)

if (!file.exists(fileTraining)) {
  download.file(fileUrlTraining, destfile = fileTraining, method ="curl")
  dateDownloaded <- date()
}

if (!file.exists(fileTesting)) {
  download.file(fileUrlTesting, destfile = fileTesting, method ="curl")
  dateDownloaded <- date()
}
```

Read each data
```{r}
pmlTraining <- read.csv(fileTraining)
pmlTesting <- read.csv(fileTesting)
```


# Exploratory Data Analysis

Check out the data
```{r, echo=FALSE}
names(pmlTraining)
names(pmlTesting)
```

We are going to predict the factor in the 160th column ‘classe’, labelled A, B, C, C, E.
```{r}
pmlTraining[,c("classe")][1:10]    #check 160th cloumn name
pmlTesting[,c("problem_id")][1:20]  #check 160th column name
#summary(pmlTraining)
#summary(pmlTraining)
```

We can also check it with compareGroups function
```{r, echo=FALSE}
library(compareGroups)
comGroup <- compareGroups(classe ~ . , data=pmlTraining)
createTable(comGroup)
```

Looking at the data we can see it contains many blank, NAs and '#DIV/0!' value.
```{r}
# Re-read data
pmlTraining <- read.csv(fileTraining, header = TRUE, na.string = c("", 'NA', "#DIV/0!"))
pmlTesting <- read.csv(fileTesting,  header = TRUE, na.string = c("", 'NA', "#DIV/0!"))
```

Select predictors, not having NA value. And remove first 7 columns bacuase it is not information from sensors 
```{r}
predictors <- colnames(pmlTraining)
predictors <- predictors[colSums(is.na(pmlTraining)) == 0]

predictors <- predictors[-(1:7)]
```

Make the final data set for analyizing
```{r}
pmlTraining <- pmlTraining[, predictors]
```


# Cross validation 
Split the pmlTraining data into two data sets : training and testing. We will allocaate 75% for taining and 25% for validating. 
```{r}
library(caret)
set.seed(2585)
inTrain<-createDataPartition(pmlTraining$classe, p=0.75, list=FALSE)
training<-pmlTraining[inTrain,]
testing<-pmlTraining[-inTrain,]
```



# Building models
## Classification tree model
```{r}
library(rpart)
modFitRpart <- train(classe ~ ., method="rpart", data=training)
predictRpart <- predict(modFitRpart, testing)
conMatRpart <- confusionMatrix(testing$classe, predictRpart)
   
conMatRpart
```
As shown above, the acccurccy of the classification tree model is vary low.


## Ramdom forest
```{r}
library(randomForest)
modFitRF <- train(classe ~.,method="rf", data=training,
              trControl = trainControl(method="cv"),
              number=3)
predictRF <- predict(modFitRF, testing)
conMatRF <- confusionMatrix(testing$classe, predictRF)
conMatRF
```


# Conclusion
From the above two confustion matrix, we can see randomForest performance is much better. 

# Project submission
```{r}
#Apply models to the 20 test cases 
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

answers <- predict(modFitRF, pmlTesting)

pml_write_files(answers)
```




# Reference 
[1] Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

[2] Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 


